---
title: "Transport Analytics"
subtitle: "Data source: Live Position Data"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    anchor_sections: yes
    code_folding: hide
    code_download: yes
    highlight: kate
    theme: journal
    number_sections: yes
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Preambles

## Load packages


```{r load-libraries}

if(!require(pacman)) install.packages("pacman")

pacman::p_load(
  arrow,
  tidymodels,
  bonsai,
  modeltime,
  tidyverse,
  stacks,
  tictoc,
  vetiver,
  pins,
  lubridate,
  plumber,
  openxlsx,
  DALEX,
  DALEXtra
)

tidymodels_prefer()
options(scipen = 999)
```


## Load dataset

```{r read-data, cache=TRUE}

tbl <- read_parquet("../IbeseLivePosition/ml_data/ml_data.parquet", as_tibble = TRUE)

```




## Data Partitioning


```{r data-split}


set.seed(1234)
trip_split <- initial_split(tbl,
                            prop = 0.75,
                            strata = "Duration")

trip_train <- trip_split %>% 
  training()

trip_test <- trip_split %>% 
  testing()
```



***



# Ensembles of Models


## Model Formula & Data preprocessing

```{r recipe-wkflw}
rec <- recipe(Duration ~ Day + Month + Hour + TimeOfDay + DistToPlant + DistCovered + TimeDiff + Altitude + Longitude + Latitude, data = trip_train)

# Now add preprocessing steps to the recipe:

trip_signed <- rec %>%
  #step_log(all_outcomes(), base = 10) %>% 
  step_impute_knn(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_spatialsign(all_numeric_predictors()) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_other(all_nominal_predictors()) %>% 
  step_filter_missing(all_nominal_predictors(), threshold = 0) %>% 
  prep()
 # step_zv(all_predictors(), skip = TRUE) %>% 
 

trip_signed

#trip_test_bake <- bake(trip_signed, new_data = trip_test)
```
***

## Resampling

```{r resampling}

set.seed(1234)

folds <- vfold_cv(trip_train, v = 5, strata = Duration)

metric <- metric_set(mape, mase, smape, rsq)

contrl_preds <- control_resamples(save_pred = TRUE, save_workflow = TRUE)

cltr_grid <- control_stack_grid()
cltr_res <- control_stack_resamples()

```

# Create Models Definition


## Lightgbm Model Specification

```{r lightgbm-model-spec}
# Lightgbm model specification

lightGBM_spec <- 
  boost_tree(
    mode = "regression",
    mtry = tune("mtry"),
    trees = tune("trees"),
    min_n = tune("min_n"),
    tree_depth = tune("tree_depth"),
    learn_rate = tune("learn_rate"),
    loss_reduction = tune("loss_reduction"),
    sample_size = tune("sample_size")
  ) %>% 
  set_engine("lightgbm")


lightGBM_wflow <- 
  workflow() %>% 
  add_model(lightGBM_spec) %>% 
  add_recipe(trip_signed)

lightGBM_wflow



set.seed(1234)

lightGBM_res <- 
  tune_grid(
    lightGBM_wflow,
    resamples = folds,
    metrics = metric,
    grid = 4,
    control = cltr_grid
  )

lightGBM_res
```




## KNN Model Specification

```{r knn-model-spec}
# knn model specification

knn_spec <- 
  nearest_neighbor(
    mode = "regression",
    neighbors = tune("k")
  ) %>% 
  set_engine("kknn")

## knn specification

knn_wflow <- 
  workflow() %>% 
  add_model(knn_spec) %>% 
  add_recipe(trip_signed)

knn_wflow



set.seed(1234)

knn_res <- 
  tune_grid(
    knn_wflow,
    resamples = folds,
    metrics = metric,
    grid = 4,
    control = cltr_grid
  )

knn_res
```







## Xgboost Model Specification

```{r xgboost-model-spec}
# xgboost model specification

xgb_spec <- 
  boost_tree(
    mtry = tune("mtry"),
    trees = tune("trees"),
    min_n = tune("min_n"),
    tree_depth = tune("tree_depth"),
    learn_rate = tune("learn_rate"),
    loss_reduction = tune("loss_reduction")
  ) %>% 
  set_engine("xgboost") %>% 
  set_mode("regression") %>% 
  translate()

xgb_wflow <- 
  workflow() %>% 
  add_model(xgb_spec) %>% 
  add_recipe(trip_signed)



set.seed(1234)

xgb_res <- 
  tune_grid(
    xgb_wflow,
    resamples = folds,
    grid = 6,
    metrics = metric,
    control = cltr_grid
  )

xgb_res
```




## Random Forest Model Specification


```{r rf-model-spec}
# Random Forest model specification

rf_mod <- rand_forest(trees = tune(),
                      min_n = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("ranger", importance='impurity')

## tune_rf_model


rf_grid <- dials::parameters(rf_mod) %>% grid_random(size = 5)

# setup workflow

rf_trip_wflow <- workflow() %>% 
  add_model(rf_mod) %>% 
  add_recipe(trip_signed)

rf_trip_wflow

## hyper-parameter tuning with cross validation

# train, test, and evaluate model performance

set.seed(1234)

rf_res <- 
  rf_trip_wflow %>% 
  tune_grid(resamples = folds,
                grid = 6,
                control = cltr_grid,
                metrics = metric) 

rf_res

```

# Stacking

## Putting together a stack

```{r stack-members}
tic("start stack models")
stacked_models <- 
  stacks() %>% 
  add_candidates(knn_res) %>% 
  add_candidates(xgb_res) %>% 
  add_candidates(rf_res) %>% 
  add_candidates(lightGBM_res) 

stacked_models %>% as_tibble()
toc()
```


## Blend the stack

```{r blend-stacks}
tic("Fit the stack models")
fit_stacked_models <- 
  stacked_models %>% 
  blend_predictions()
toc()
```

## Plots

```{r plots}
theme_set(theme_bw())

autoplot(fit_stacked_models)

autoplot(fit_stacked_models, type = "members")


```



If these results were not good enough, `blend_predictions()` could be called again with different values of penalty. As it is, `blend_predictions()` picks the penalty parameter with the numerically optimal results. To see the top results:

```{r autoplot}
autoplot(fit_stacked_models, type = "weights")
```


Now that we know how to combine our model output, we can fit the candidates with non-zero stacking coefficients on the full training set.

## Fit Stacked Models

```{r fit-stacked}
tic("fit the model members")
fit_stacked_models<- 
  fit_stacked_models %>% 
  fit_members()
toc()
```


To identify which model configurations were assigned what stacking coefficients, we can make use of the `collect_parameters()` function:


```{r par-knn-res}
collect_parameters(fit_stacked_models, "knn_res")
```


```{r par-xgb-res}
collect_parameters(fit_stacked_models, "xgb_res")
```


```{r par-rf-res}
collect_parameters(fit_stacked_models, "rf_res")
```


```{r}
collect_parameters(fit_stacked_models, "rf_res")
```

## Predict with the Stacked Models

This object is now ready to predict with new data!

```{r predict-stacked-models}
test_model <- 
  trip_test %>% 
  bind_cols(predict(fit_stacked_models, .))

test_df <- test_model %>% 
  mutate(Predicted = DateTimeReceived + seconds(.pred), .after = Arrival) %>%
  select(TripID, Reference, Arrival, Predicted) %>% 
  arrange(TripID, Reference, Arrival, Predicted) %>% 
  group_by(TripID, Reference, Arrival) %>% 
  slice_tail(n = 1) %>% 
   ungroup() %>% 
   mutate(Variance = paste0(round((as.numeric(Arrival - Predicted))/3600), " hrs"))

test_df
```

```{r write-test-df, eval=FALSE, echo=FALSE}

hs <- createStyle(
  textDecoration = "BOLD", fontColour = "#FFFFFF", fontSize = 12,
  fontName = "Arial Narrow", fgFill = "#4F80BD"
)

write.xlsx(test_df,
           file = "Actual_vs_Prediction.xlsx",
           colNames = TRUE, borders = "rows", headerStyle = hs
)

```



Juxtaposing the predictions with the true data:

```{r plot-predictions}
ggplot(test_model) +
  aes(x = Duration,
      y = .pred) +
  geom_point() +
  coord_obs_pred()
```


How do the stacks predictions perform, though, as compared to the members’ predictions? We can use the `type = "members"` argument to generate predictions from each of the ensemble members.


```{r compare-members}
member_preds <- 
  test_model %>% 
  select(Duration) %>% 
  bind_cols(predict(fit_stacked_models, test_model, members = TRUE)) %>% 
  rename(stackedModels = `.pred`)

member_preds
```


Now, evaluating the root mean squared error from each model:

```{r rmse-comparison}
map_dfr(member_preds, rmse, truth = Duration, data = member_preds) %>%
  mutate(member = colnames(member_preds))
```



and the $r^2$ from each model:

```{r rf-comparison}
map_dfr(member_preds, rsq, truth = Duration, data = member_preds) %>%
  mutate(member = colnames(member_preds))
```



# Model Versioning & Deployment


```{r}
class(fit_stacked_models)
```

```{r}
fit_stacked_models
```




```{r vetiver-model, eval=TRUE}
v <- vetiver_model(fit_stacked_models, "dcp_ibese_truck_arrival")
v
```


```{r set-up-pins, eval=TRUE}
model_board <- board_folder("pins-r", versioned = TRUE)
model_board %>% vetiver_pin_write(v)
model_board %>% pin_versions("dcp_ibese_truck_arrival")
vetiver_write_plumber(model_board, "dcp_ibese_truck_arrival", rsconnect = FALSE)
```

## Create a REST API for deployment

```{r eval=FALSE}
pr() %>% 
  vetiver_api(v) %>% 
  pr_run(port = 9090)
```


## Predict from the model endpoint

```{r eval=FALSE}
endpoint <- vetiver_endpoint("http://127.0.0.1:9090/predict")
```


## Generate a Dockerfile

```{r eval=FALSE}
vetiver_write_docker(v)
```


# Session Information

```{r session-info}
sessionInfo()
```

# Variable Importance

```{r explainer}

explain_stack <- explain_tidymodels(fit_stacked_models, data = trip_test %>% select(-Duration, -TripID, -Reference, -Geofence, -DateTimeReceived, -Arrival), y = trip_test$Duration,
                                    verbose = FALSE, label ="Stack Model")

```




```{r ggplot_imp_function}
ggplot_imp <- function(...) {
  obj <- list(...)
  metric_name <- attr(obj[[1]], "loss_name")
  metric_lab <- paste(metric_name, 
                      "after permutations\n(higher indicates more important)")
  
  full_vip <- bind_rows(obj) %>%
    filter(variable != "_baseline_")
  
  perm_vals <- full_vip %>% 
    filter(variable == "_full_model_") %>% 
    group_by(label) %>% 
    summarise(dropout_loss = mean(dropout_loss))
  
  p <- full_vip %>%
    filter(variable != "_full_model_") %>% 
    mutate(variable = fct_reorder(variable, dropout_loss)) %>%
    ggplot(aes(dropout_loss, variable)) 
  if(length(obj) > 1) {
    p <- p + 
      facet_wrap(vars(label)) +
      geom_vline(data = perm_vals, aes(xintercept = dropout_loss, color = label),
                 size = 1.4, lty = 2, alpha = 0.7) +
      geom_boxplot(aes(color = label, fill = label), alpha = 0.2)
  } else {
    p <- p + 
      geom_vline(data = perm_vals, aes(xintercept = dropout_loss),
                 size = 1.4, lty = 2, alpha = 0.7) +
      geom_boxplot(fill = "#91CBD765", alpha = 0.4)
    
  }
  p +
    theme(legend.position = "none") +
    labs(x = metric_lab, 
         y = NULL,  fill = NULL,  color = NULL)
}
```



```{r var-imp-plot, fig.cap="Global explainer for the stacked ensemble tidymodel on the trip data", fig.width = 9, fig.height=5}

set.seed(1234)

  explain_stack %>% 
  model_parts() %>% 
  ggplot_imp()

```

# Daily Prediction

* Predict the arrival of inbound trucks that are yet to arrive within 24 hrs period.

```{r predict-data-prep}

trip_to_predict_tbl <- read_parquet("../IbeseLivePosition/ml_data/trips_to_predict.parquet", as_tibble = TRUE)
```



```{r predict-inbound-trucks}

predict_df <- trip_to_predict_tbl %>% 
  bind_cols(predict(fit_stacked_models, .)) %>% 
  mutate(Predicted = DateTimeReceived + seconds(.pred), .after = Reference) %>%
  select(TripID, Reference, Predicted) %>% 
  group_by(TripID, Reference) %>% 
  slice_tail(n = 1) %>% 
  ungroup() %>% 
  filter(as.Date(Predicted) == today()) %>% 
  .[order(.$Predicted),]

predict_df
```




```{r write-predict-df, eval=FALSE, echo=FALSE}

hs <- createStyle(
  textDecoration = "BOLD", fontColour = "#FFFFFF", fontSize = 12,
  fontName = "Arial Narrow", fgFill = "#4F80BD"
)

write.xlsx(predict_df,
           file = "Ibese Daily Inbound Trucks Arrival Prediction-0916.xlsx",
           colNames = TRUE, borders = "rows", headerStyle = hs
)

```



```{r prediction-vs-actual, eval=FALSE, echo=FALSE, include=FALSE}

library(readxl)

file <- read_excel("Ibese Daily Inbound Trucks Arrival Prediction-0916.xlsx")

confirmation <- file %>% 
  inner_join(Actual_tbl) %>% 
  select(-Time) %>% 
  mutate(Variance = round(difftime(Arrival, Predicted, units = "hours"),2))

confirmation


data_list <- list("prediction" = file,
                  "confirmation" = confirmation)

hs <- createStyle(
  textDecoration = "BOLD", fontColour = "#FFFFFF", fontSize = 12,
  fontName = "Arial Narrow", fgFill = "#4F80BD"
)

write.xlsx(data_list,
           file = "Ibese Daily Inbound Trucks Arrival PredictionVSConfirmation-0916.xlsx",
           colNames = TRUE, borders = "rows", headerStyle = hs,
           asTable = TRUE, withFilter = TRUE
)



```

