---
title: "Trips with zero TripIDs"
subtitle: "Trips with zero TripIDs Statistical Analysis"
author: "Olumide Oyalola"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    anchor_sections: yes
    code_folding: show
    code_download: yes
    highlight: kate
    theme: journal
    number_sections: yes
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Packages

```{r load-packages}
# Install pacman package if needed
if(!require("pacman")) install.packages("pacman")

# load the required packages

pacman::p_load(
  httr,
  jsonlite,
  tidyjson,
  tidyverse,
  lubridate,
  geosphere,
  anytime,
  tictoc,
  stringi,
  maptools,
  geosphere,
  sf,
  sp,
  openxlsx,
  leaflet,
  magrittr,
  janitor,
  collapse,
  DT
)
```

# Load Data

```{r load-data}
trpID_is_zero <- read_rds("tripIDisZero.RDS")
```




# Summary

* Two weeks long hourly data starting from `2022-06-10 08:00:08` to `2022-06-25 00:00:15`. The API was pinged hourly within the closed timestamp above. All trips with TripIDs equal zero were filtered and analyzed. The analysis is as detailed in the subsections below.

* Below are the insights derived from the filtered data (Trips with `TripIDs` equal zero within the stated timestamp above).

  1. 452,229 out of 1,001,595 is the total number of live position data corresponding to trips with `TripID` equal zero. This corresponds to 45% of the live position data captured within the time frame having `TripID` equal zero.
  
  2. Trips within this period have 194 unique `Geofences`. Below are samples of the distinct `Geofences`.
  
  `r trpID_is_zero %>% 
    pull(Geofence) %>% 
    unique() %>% 
    .[160:194]`
    
  
  3. `HGV`, `FuelTanker` and `MixerTruck` are the top 3 assets with no TripID.
  
  4. * Within the two weeks time window, No `InService` truck within the Ibese geofences (Ibese plant geofence and several other geofences within it) has zero `TripID`.



# Analysis



## Data Description

* Summary of the `numeric` and `character` type variables.

```{r describe-data, comment=""}

trpID_is_zero %>% 
select_if(function(col) is.numeric(col) | 
                        is.character(col)) %>% 
descr() 

```


```{r assetClass}
trpID_is_zero %>%
  mutate(AssetName = case_when(AssetClass == 0 ~ "Car",
                               AssetClass == 1 ~ "Van",
                               AssetClass == 2 ~ "HGV",
                               AssetClass == 3 ~ "PSV",
                               AssetClass == 4 ~ "Motor_Cycle",
                               AssetClass == 5 ~ "Ambulance",
                               AssetClass == 6 ~ "_4X4",
                               AssetClass == 7 ~ "Container",
                               AssetClass == 8 ~ "Trailer",
                               AssetClass == 9 ~ "Crane",
                               AssetClass == 10 ~ "Dump_Truck",
                               AssetClass == 11 ~ "Fire_Engine",
                               AssetClass == 12 ~ "Train",
                               AssetClass == 13 ~ "Mini_Bus",
                               AssetClass == 14 ~ "Helicopter",
                               AssetClass == 15 ~ "Plane",
                               AssetClass == 16 ~ "Police_Car",
                               AssetClass == 17 ~ "Trawler",
                               AssetClass == 18 ~ "camper",
                               AssetClass == 19 ~ "Speed_Boat",
                               AssetClass == 20 ~ "Ship",
                               AssetClass == 21 ~ "Farm_Vehicle",
                               AssetClass == 22 ~ "Generator",
                               AssetClass == 23 ~ "Mobile_Tower",
                               AssetClass == 24 ~ "Taxi",
                               AssetClass == 25 ~ "Dog",
                               AssetClass == 26 ~ "Cat",
                               AssetClass == 27 ~ "Pet_Other",
                               AssetClass == 28 ~ "Human",
                               AssetClass == 29 ~ "Other_Vehicle",
                               AssetClass == 30 ~ "Custom_Image",
                               AssetClass == 31 ~ "Forklift",
                               AssetClass == 32 ~ "Construction_Equipment",
                               AssetClass == 33 ~ "Bicycle",
                               AssetClass == 34 ~ "CementTruck",
                               AssetClass == 35 ~ "CoalTruck",
                               AssetClass == 36 ~ "RecoveryTruck",
                               AssetClass == 37 ~ "TruckMountedCrane",
                               AssetClass == 38 ~ "TowTruck",
                               AssetClass == 39 ~ "FuelTanker",
                               AssetClass == 40 ~ "MixerTruck",
                               AssetClass == 41 ~ "QuadBike",
                               AssetClass == 42 ~ "Truck",
                               AssetClass == 43 ~ "IOT",
                               AssetClass == 44 ~ "Truck_Head",
                               AssetClass == 45 ~ "Backhoe_Loader",
                               AssetClass == 46 ~ "Dozer",
                               AssetClass == 47 ~ "Excavator",
                               AssetClass == 48 ~ "Motor_Grader",
                               AssetClass == 49 ~ "Polaris_Ranger",
                               AssetClass == 50 ~ "Sheep_Foot_Roller",
                               AssetClass == 51 ~ "Vibratory_Roller",
                               AssetClass == 52 ~ "Wheel_Loader")) %>% 
  count(AssetClass, AssetName) %>% 
  arrange(n) %>% 
  rename(Frequency = n) %>% 
  datatable(caption = "Distribution of Trips with zero tripIDs by AssetClass") %>% 
  formatRound(columns = c('Frequency'), digits = 0)
```

## Cement Trucks outside of Ibese Geofence with no TripIDs

```{r}

ibese_geofence <- c("Ibese", "IBESE")


trpID_is_zero %>% 
  mutate(Geofence = case_when(Geofence == "" ~ "NA",
                              TRUE ~ Geofence)) %>% 
  filter(!str_detect(Geofence, paste(ibese_geofence, collapse = "|"))) %>% 
  filter(AssetClass == 34) %>% 
  count(Reference, Geofence) %>% 
  rename(Count = n) %>% 
  datatable(caption = "Cement Trucks outside of Ibese Geofence with no TripIDs") %>% 
  formatRound(columns = c('Count'), digits = 0)
```

# Efficiency of Utilization

* Average time it takes before an operational truck is assigned to a trip

```{r read-wks-all-livePosition-data}
two_wks_all_position_tbl <- read_rds("2wks-All-LivePosition.RDS")
```



```{r}
geo_proj = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"


position <- two_wks_all_position_tbl %$% 
  matrix(c(Longitude, Latitude), ncol = 2)

position_sf <- SpatialPoints(position, 
                              proj4string=CRS(geo_proj)) %>% 
   st_as_sf()
 
plant <- matrix(c(3.05120587348938, 7.00300851960855), ncol = 2)

ibese_sp <- SpatialPoints(plant, proj4string=CRS(geo_proj)) %>% 
  st_as_sf()

plant_buffer <- st_buffer(ibese_sp, 5000)

 
row_id <- st_intersects(plant_buffer, position_sf)[[1]]
 
check_df <- two_wks_all_position_tbl[row_id,]

#inservice - tripid == 0; the tripid changes from zero
# the truck is expected to be within the geofence of the plant.
# use 2km buffer
# the time elapse from when the tripID=0 to when it turned to another value except zero.




```


## Cement trucks Utilization efficiency check

```{r efficiency}

check_df %>% 
  filter(AssetStatus == "InService", AssetStatus %in% c(34, 42)) %>% 
  select(DateTimeLocal, Reference, TripID, Geofence, AssetClass) %>% 
  arrange(DateTimeLocal, Reference, AssetClass, TripID) %>% 
  group_by(Reference, AssetClass, TripID) %>% 
  slice(c(1, n())) %>% 
  mutate(`Delay (Hrs)` = case_when((TripID == 0 & lag(TripID) == 0) | (TripID == 0 & lag(TripID != 0)) ~ as.numeric(difftime(DateTimeLocal, 
                                        lag(DateTimeLocal, 
                                            default = first(DateTimeLocal)), 
                                        units = "hours")))) %>% 
  filter(!is.na(`Delay (Hrs)`)) %>% 
  datatable(caption = "Cement trucks Utilization efficiency check") %>% 
  formatRound(columns = c('Delay (Hrs)'), digits = 0)
  

```

* Within the two weeks time window, No `InService` truck within the Ibese geofences (Ibese plant geofence and several other geofences within it) has zero `TripID`.



## Other `assetclass` utilization efficiency check

```{r efficiency2}

check_df %>% 
  filter(AssetStatus == "InService", !AssetStatus %in% c(34, 42)) %>% 
  select(DateTimeLocal, Reference, TripID, Geofence, AssetClass) %>% 
  arrange(DateTimeLocal, Reference, AssetClass, TripID) %>% 
  group_by(Reference, AssetClass, TripID) %>% 
  slice(c(1, n())) %>% 
  mutate(`Delay (Hrs)` = case_when((TripID == 0 & lag(TripID) == 0) | (TripID == 0 & lag(TripID != 0)) ~ as.numeric(difftime(DateTimeLocal, 
                                        lag(DateTimeLocal, 
                                            default = first(DateTimeLocal)), 
                                        units = "hours")))) %>% 
  filter(!is.na(`Delay (Hrs)`)) %>% 
  datatable(caption = "Other assetclass utilization efficiency check") %>% 
  formatRound(columns = c('Delay (Hrs)'), digits = 0)
  

```

